---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
#### Load the data
```{r, results='markup', warning=TRUE, message=TRUE}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
activityData <- read.csv('activity.csv')
```
## What is mean total number of steps taken per day?
stepsByDay <- tapply(activityData$steps, activityData$date, sum, na.rm=TRUE)

### 1. Make a histogram of the total number of steps taken each day
```{r}
qplot(stepsByDay, xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)

### 2. Calculate and report the mean and median total number of steps taken per day
```{r}
####Calculate the mean 
stepsByDayMean <- mean(stepsByDay)

####Calculate the median
stepsByDayMedian <- median(stepsByDay)

```
## What is the average daily activity pattern?
```{r}
averageStepsPerTimeBlock <- aggregate(x=list(meanSteps=activityData$steps), by=list(interval=activityData$interval), FUN=mean, na.rm=TRUE)
```
### 1. Make a time series plot (i.e. \color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
ggplot(data=averageStepsPerTimeBlock, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken") 

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
